Incremental Forward Reasoning for White-Box~Proof~Search

### 2025-02-13 paper.tex
Some clarifications in abstract

\begin{abstract}
  Several proof assistants provide automation tactics based on tableau-style tree search, such as Isabelle's and Rocq's auto and Lean's Aesop.
  We consider \emph{forward rules}, which apply a given theorem, say, $A → B → C$, to any goal containing hypotheses $A$ and $B$, adding $C$ as a hypothesis.
  When treated naively, such rules are tried on every goal encountered during the search, leading to repeated unifications of premises $A$ and $B$ with the hypotheses of each goal.
  We present an approach to forward rules that takes advantage of similarities between successive goals.
  For each goal, we cache partial applications of forward rules in a custom data structure, the \emph{partial match state}, which enables efficient updates.
  Our technique is compatible with any search strategy and most logics.
  It has been implemented in Aesop.

@@

% \begin{abstract}
%   Many proof assistants provide automation tactics based on simple, tableau-style tree search, such as Isabelle's and Rocq's auto or Lean's Aesop.
%   These tactics solve a given goal by recursively applying \emph{rules}, which return zero or more subgoals.
%   In most cases, the rules themselves can be arbitrary tactics, so they can return any set of subgoals corresponding to an admissible inference in the prover's logic.

%   In this setting, we consider \emph{forward rules}, which apply a given theorem, say $A → B → C$, to any goal containing hypotheses $A$ and $B$, adding $C$ as an additional hypothesis.
%   When treated naively, such rules are tried on every goal encountered during the search, leading to repeated unifications of premises $A$ and $B$ with the hypotheses of each goal.
%   Much of this work is redundant since goals often have similar hypotheses.

%   We present an implementation of forward rules that takes advantage of these similarities.
%   For each goal, we cache certain partial applications of forward rules in a custom data structure, the \emph{partial match state}.
%   This enables efficient updates, so if goals have similar hypotheses, we can quickly update the partial match state rather than re-running all forward rules.

%   Our technique is compatible with any search strategy and most logics.
%   It has been implemented in the Aesop proof search tactic for Lean.

%   \keywords{Forward Reasoning \and Forward Chaining \and Tactics \and Interactive Theorem Proving.}
% \end{abstract}

>>>>

\begin{abstract}
  Several proof assistants provide automation tactics based on tableau-style tree search, such as Isabelle's and Rocq's auto and Lean's Aesop.
  In this setting we consider \emph{forward rules}, which apply a given theorem, say, $A → B → C$, to any goal containing hypotheses $A$ and $B$, adding $C$ as a new hypothesis.
  When treated naively, such rules are tried on every goal encountered during the search, leading to repeated unifications of premises $A$ and $B$ with the hypotheses of each goal.
  We present an approach to forward rules that avoids some of this repeated work by taking advantage of similarities between successive goals.
  For each goal, we cache partial applications of forward rules in a custom data structure, the \emph{partial match state}, which enables efficient updates.
  Our technique is compatible with any search strategy and most logics.
  It has been implemented in Aesop.

@@
